P259: Gene set enrichment analysis (GSEA)

### 2021-05-19 3.P259.pDC_GSEA.Rmd
Add DEG enrichment analysis

```{r echo=FALSE}
source("scripts/GSEA.score.plot.R")
```

#### Hallmark FDR < 0.1

```{r echo=FALSE, fig.height=5, fig.width=8.5}
gsea.plot(gsea = h_GSEA.result, fdr.cut = 0.1, 
          prefix = "H", print.plot = TRUE, save.plot = FALSE)
```

#### Curated gene sets (C2) FDR < 0.05

```{r echo=FALSE, fig.height=25, fig.width=8.5}
gsea.plot(gsea = c2_GSEA.result, fdr.cut = 0.05, 
          prefix = "C2", print.plot = TRUE, save.plot = FALSE)
```

#### GO gene sets (C5) FDR < 0.05

```{r echo=FALSE, fig.height=25, fig.width=8.5}
gsea.plot(gsea = c5_GSEA.result, fdr.cut = 0.05, 
          prefix = "C5", print.plot = TRUE, save.plot = FALSE)
```

>>>>

# Hypergeometric enrichment of DEGs

```{r echo=FALSE}
fdr.cut <- 0.1
```

Test for Broad gene set enrichment in differentially expressed genes (DEG) as defined by change with virus (*e.g.* significant for virus in untreated and/or treated donors) AND different between untreated and treated donors in media OR RV. DEGs are defined at FDR < `r fdr.cut`

```{r}
#Define change with virus
DEG1.v <- read_csv("results/gene_level/P259.1_gene_pval.csv") %>% 
  filter(group %in% c("none_HRV - none_none", "EOS.supp_HRV - EOS.supp_none") &
           adj.P.Val <= fdr.cut)
#Define change with treament
DEG1.t <- read_csv("results/gene_level/P259.1_gene_pval.csv") %>% 
  filter(group %in% c("EOS.supp_none - none_none", "EOS.supp_HRV - none_HRV") &
           adj.P.Val <= fdr.cut)

#Define change with virus
DEG2.v <- read_csv("results/gene_level/P259.2_gene_pval.csv") %>% 
  filter(group %in% c("none_HRV - none_none", "AntiIL5_HRV - AntiIL5_none") &
           adj.P.Val <= fdr.cut)
#Define change with treament
DEG2.t <- read_csv("results/gene_level/P259.2_gene_pval.csv") %>% 
  filter(group %in% c("AntiIL5_none - none_none", "AntiIL5_HRV - none_HRV") &
           adj.P.Val <= fdr.cut)

#Format to ENTREZ

#Format in list
DEG.ls <- list()
DEG.ls[["EOSsup"]] <- intersect(DEG1.v$hgnc_symbol, DEG1.t$hgnc_symbol)
DEG.ls[["AntiIL5"]] <- intersect(DEG2.v$hgnc_symbol, DEG2.t$hgnc_symbol)
```

```{r message=FALSE, warning=FALSE}
#Script for running term enrichment
source("https://raw.githubusercontent.com/kdillmcfarland/R_bioinformatic_scripts/master/hypergeo_enricher.R")
```

## Run enrichment
#### Hallmark (H)

```{r h.2, warnings=FALSE, message=FALSE, eval=FALSE}
enrich.fxn(gene.list = DEG.ls, ID.type="SYMBOL",
           category = "H",
           genome = "org.Hs.eg.db", 
           basename = "DEG",
           outdir = "results/enrichment/")
```

```{r c2.2, warnings=FALSE, message=FALSE, eval=FALSE}
enrich.fxn(gene.list = DEG.ls, ID.type="SYMBOL",
           category = "C2", subcategory="CP",
           genome = "org.Hs.eg.db", 
           basename = "DEG",
           outdir = "results/enrichment/")
```

```{r c5.2, warnings=FALSE, message=FALSE, eval=FALSE}
enrich.fxn(gene.list = DEG.ls, ID.type="SYMBOL",
           category = "C5", subcategory = "GO:BP",
           genome = "org.Hs.eg.db", 
           basename = "DEG",
           outdir = "results/enrichment/")
```

## Significant enrichment

```{r echo=FALSE, message=FALSE}
fdr.cut <- 0.05
#Load results for faster knitting
h_enrich.result <- read_csv("results/enrichment/enrich_DEG_H.csv")
c2_enrich.result <- read_csv("results/enrichment/enrich_DEG_C2_CP.csv")
c5_enrich.result <- read_csv("results/enrichment/enrich_DEG_C5_GO.BP.csv")

#Combine and filter signif
enrich.signif <- bind_rows(h_enrich.result,c2_enrich.result,c5_enrich.result)%>% 
  filter(p.adjust <= fdr.cut) 
```

Gene sets enriched in DEGs at FDR < `r fdr.cut`

```{R echo=FALSE}
enrich.signif %>% 
  separate(Description, into=c("set","pathway"), sep="_", extra = "merge") %>% 
  dplyr::select(group, set, pathway, p.adjust, SYMBOLs) %>% 
  arrange(group, set, p.adjust) %>% 
  kable() %>%
kable_styling(bootstrap_options = "striped", full_width = FALSE) %>% 
  collapse_rows(1:2, valign = "top")
```
